---
title: "P8105_HW2"
author: "Meng Fang"
date: '2022-09-19'
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(tidyr)
```


## Problem 1

- Data Cleaning

```{r}
nyc_transit <- read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% janitor::clean_names()
```

```{r}
tidy_transit <- nyc_transit %>%
  select(2:18, 20, 23) %>%
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

tidy_transit
```

the dataset contains `r nrow(tidy_transit)` rows and `r ncol(tidy_transit)` columns. It contains variables: `r names(tidy_transit)`. The above cleaning processes include: select desirable variables, change variables: route8, rout9, route10 into character. and convert the entry variable from character to logical variable.

- number of distinct station

```{r}
tidy_transit %>%
  distinct(line, station_name) %>%
  nrow()
```

There are 465 distinct stations.

- number of stations that are ada compliant

```{r}
tidy_transit %>%
  filter(ada == TRUE) %>%
  distinct(line, station_name) %>%
  nrow()
```

84 stations are ada compliant

- proportion of station entrances / exits without vending 

```{r}
tidy_transit %>%
  filter(vending == "NO") %>%
  pull(entry)%>%
  mean()
```

There are 37.7% of the stations entrances / exits without vending.

- Number of stations serving A

```{r}
tidy_transit %>% 
  pivot_longer(route1:route11, names_to = "route_num", values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct() %>%
  nrow()
```

60 stations serving line A

- Number of stations serving A that are ADA compliant

```{r}
tidy_transit %>% 
  pivot_longer(route1:route11, names_to = "route_num", values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct() %>%
  nrow()
```

17 serving A that are ADA compliant

## Problem 2

(1) Clean Mr.Trash Wheel Data 

```{r}
mr_tw <- read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N550") %>% janitor::clean_names()
```



```{r}
tidy_mrtw <- mr_tw %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(vessel_type = "Mr.Trash Wheel") %>%
  mutate(year = as.integer(year), dumpster = as.integer(dumpster))%>%
  drop_na(dumpster)

tidy_mrtw
```

(2) Clean Professor Trash Wheel Data 

```{r}
prof_tw <- read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M97") %>% janitor::clean_names()
```


```{r}
tidy_proftw <- prof_tw %>%
  drop_na(dumpster) %>%
  mutate(year = as.integer(year), vessel_type = "Professor Trash Wheel") 

tidy_proftw
```

(3) Bind 2 Data Together

As the professor trash wheel sheet doesn't have the `sports_balls` variable, we need to add a `sports_balls` variable with all `NA`s, indicating that we don't know the the number of balls collected by professor trash wheel.

```{r}
tidy_proftw2 <- tidy_proftw %>% mutate(sports_balls = as.integer(NA))
tidy_tw <- bind_rows(tidy_mrtw, tidy_proftw2)
tidy_tw
```


Combining those two datasets, we have a combined dataset that contains `r nrow(tidy_tw)` rows and `r ncol(tidy_tw)` columns. There are in total `r nrow(tidy_tw)` obervations in the combined dataset. An added `vessel_type` variable specifies the vessel that the observation belongs to. some other key variables include: dumpster number, month, year, date, weight, volume, different kinds of trash and their amount, homes powered by trash-made electricity. Some values for certain variables are missing, in this case, we are keeping those missing observations as it will still provide some useful information on the associated date.


- total weight of trash collected by professor trash wheel

```{r}
tidy_tw %>%
  filter(vessel_type == "Professor Trash Wheel") %>%
  select(weight_tons) %>%
  sum()
```

- total number of sports balls collected by Mr. Trash Wheel in 2020

```{r}
tidy_tw %>%
  filter(year == 2020, vessel_type == "Mr.Trash Wheel") %>%
  select(sports_balls) %>%
  sum()
```

## Problem 3

(1) Clean pols-month.csv

```{r}
pols_month <- read_csv(file = "./pols-month.csv") %>% janitor::clean_names()
```

First, we check whether in certain period of time there are more than one president. 

```{r}
pols_month %>% filter(prez_gop != 0 & prez_gop != 1)
```

```{r}
pols_month %>% filter(prez_dem != 0 & prez_dem != 1)
```

We notice that in from August to December in 1974, there are two presidents that are republicans shown in the dataset. In reality, Ford took Nixon's presidency in August 1974 and Nixon resigned at that time, some we just consider in these periods we have a republican president.

```{r}
tidy_pm <- pols_month %>%
  separate(mon, into = c("year","month","day"), sep = "-") %>%
  mutate(month = month.name[as.numeric(month)], year = as.integer(year)) %>%
  mutate(president = ifelse(prez_gop >= 1, "Republican", "Democrat")) %>%
  select(-prez_gop,-prez_dem,-day)
```

(2) Second, clean the data in snp.csv using a similar process

```{r}
snp <- read_csv(file = "./snp.csv") %>% janitor::clean_names()
```



```{r}
tidy_snp <- snp %>%
  separate(date, into = c('month', 'day', 'year'), sep = '/') %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>%
  mutate(year = as.integer(ifelse(year <= 15, year + 2000, year + 1900))) %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  select(year, month, close)

```


(3) Third, tidy the unemployment data so that it can be merged with the previous datasets

```{r}
unemploy <- read_csv(file = "./unemployment.csv") %>% janitor::clean_names()
```

```{r}
tidy_unemploy <- unemploy %>%
  pivot_longer(2:13, names_to = "month", values_to = "unemploy_perc") %>%
  mutate(month = recode(month,
                        'jan' = 'January', 
                        'feb' = 'February', 
                        'mar' = 'March', 
                        'apr' = 'April', 
                        'may' = 'May', 
                        'jun' = 'June', 
                        'jul' = 'July', 
                        'aug' = 'August', 
                        'sep' = "September", 
                        'oct' = 'October', 
                        nov = 'November', 
                        dec = 'December'), 
         year = as.integer(year))

```

(4) Combine all three dataset together

```{r}
combined_df <-  merge(tidy_pm, tidy_snp) %>% merge(tidy_unemploy)

```

Combining those three datasets, and dropping missing values for each variable, we have a combined dataset that contains `r nrow(combined_df)` rows and `r ncol(combined_df)` columns. There are in total `r nrow(combined_df)` in the combined dataset., the range of years is: `r max(combined_df$year)-min(combined_df$year)`. Some key variables include:

- year
- month
- gov_gop: the number of republican governors on the associated date
- sen_gop: the number of republican senators on the associated date
- rep_gop: the number of republican representatives on the associated date
- gov_dem: the number of democratic governors on the associated date
- sen_dem: the number of democratic senators on the associated date
- rep_dem: the number of democratic representatives on the associated
- president: the party of the president on the associated date
- close: the closing values of the S&P stock index on the associated date
- unemploy_perc: the unemployment percentage on the associated date
